---
title: "Assignment 7"
author: "Anastasiya Pautarak"
format: html
editor: visual
---

# **Analysis of the Iris Dataset using tidyverse**

This report demonstrates how to use tidyverse for reshaping and analyzing the iris dataset.

## Libraries:

Load the tidyverse package which includes dplyr, tidyr, readr, ggplot2, ..., etc.

```{r}
library(tidyverse)

```

## Task 1: Transform the dataset into long format:

Reshape the data so that each measurement trait becomes a row instead of a column.

1.  Load iris dataset and pivot to long format

2.  View the reshaped data : `head(iris_long)`

3.  Now we will have three columns in the table: Species, Trait, Size

```{r}
iris_long <- iris %>%
  pivot_longer(cols = -Species, # transform all columns except Species
               names_to = "Trait",
               values_to = "Size")
head(iris_long)

```

## Task 2.1: Mean length and width for each species and trait

-   To calculate the length or width for each trait within each species.

```{r}
iris_long %>%
  group_by(Species, Trait) %>%
  summarize(MeanSize = mean(Size), .groups = "drop")

```

## Task 2.2: Largest and widest petals and sepals for *virginica* and *versicolor*

-   filter for virginica and versicolor species and find the maximum size per trait

-   keep only the two species

```{r}
iris_long %>%
  filter(Species %in% c("virginica", "versicolor")) %>%
  group_by(Species, Trait) %>%
  summarize(MaxSize = max(Size), .groups = "drop")
```

## Task 2.3: Count how many observations are above the mean

1.  Step 1: Add the group mean as a new column

```{r}
iris_above_mean <- iris_long %>%
  group_by(Species, Trait) %>%
  mutate(MeanSize = mean(Size),
         Above_Mean = Size > MeanSize)
```

2.  Step 2: Count how many are above mean

-   Join mean values to original data
-   Create a new column(T/F)

```{r}
iris_above_mean %>%
  group_by(Species, Trait) %>%
  summarize(Count_Above_Mean = sum(Above_Mean), .groups = "drop")
```
